SHIV: Reducing Supervisor Burden using Support Vectors to Efficiently Learn Robot Control Policies from Demonstrations

### 2015-08-21 paper.tex
added figs and stuff

\begin{abstract}
Algorithms such as DAgger for robot learning from demonstrations can be very effective for problems where system dynamics and reward functions are unknown, but can impose a burden on supervisors to provide labels (control signals) as states are encountered during policy rollouts.   To reduce this burden we introduce a new query selection method for stream-based active learning. The SHIV algorithm (Svm-based reduction in Human InterVention) only requests supervision for risky stater.  To facilitate scaling and outlier rejection, risk is based on distance to an approximate level set boundary defined by a One Class support vector machine.  We compare SHIV with DAgger in three policy learning experiments: 1) a driving simulator, 2) multi-joint robot arm grasping in clutter simulation with Box2D, and 3) surgical needle insertion.  Results suggest that SHIV can learn policies and reduce operator burden significantly  (by a factor of 3x to 5x).

%In traditional imitation learning, a human expert provides a few demonstrations to the robot, which then \emph{offline} learns a policy to extrapolate these examples to new situations. One challenge is that during execution of the learned policy, a small but inevitable error will amplify, taking the robot further and further away from the regions of state space for which it has seen examples.

%\emph{Online} imitation learning algorithms such as DAgger address this challenge by iteratively executing a trained policy, and asking the expert for additional examples along the states visited by that policy. Such algorithms perform well in new situations, but rely much more heavily on extensive human supervision.

%Our goal is to reduce the amount of human supervision required in imitation learning, without sacrificing performance. \emph{Our key insight is that the robot should ask for supervision only when its current trained policy has high risk.}

%We introduce a method for online imitation learning that evaluates risk based on novelty of a state, as well as prior difficulty in labeling a nearby states. Our algorithm, SHIV (SVM-Based Reduction in Human Intervention), follows the trained policy when evaluated risk is low, and asks for human supervision otherwise.

@@

\begin{figure}[t!]
\centering
\includegraphics[width=12cm, height=6cm]{figures/teaser.eps}
\caption{ 
Given a set of states corresponding to images of a race-track simulation (visualized in $\mathbb{R}^2$) with expert
control inputs, we consider a quantile super levelset estimate (shaded) as an approximation of the riskiness of
executing the trained policy at a given state. By varying the quantile threshold, we can increase or decrease the size
of the quantile super-levelset area. In our approach, points outside the super-levelset are considered risky and, during
training, roll-outs of the policy that leave the super-levelset require control demonstrations from the supervisor,
while roll-outs within the super-levelset can be performed autonomously. 
}
\vspace*{-10pt}
\label{fig:dis_traveled}
\end{figure}

@@

\section{Problem Statement}
\subsection{Notation and Background}
{\color{blue} [-- TODO: SENSOR MODEL --]}
We denote by $\mathcal{X}$ the set consisting of \emph{observable states} for a robot task, consisting, for example of
high-dimensional vectors corresponding to images from a camera, or joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of \emph{allowable control inputs} for the robot, which can be discrete or continuous in
nature. The environment is
modeled to have dynamics that are stationary and Markovian, such that the the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in \mathcal{U}$,
so that $p(\bx_{N+1}|\bu_{N},\bx_{N}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{N+1}|\bu_{N}, \bx_N)$.  
When no confusion arises, we denote the probability density over the initial state also by $p:\mathcal{X}\to
\mathbb{R}$. By a trajectory, we mean a finite series of pairs of $T$ states visited and corresponding
control inputs at these states $\tau = (\mathbf{x}_0,\mathbf{u}_1, ...., \mathbf{x}_T,\mathbf{u}_T)$, where $\bx_i\in \mathcal{X}$
and $\bu_i\in \mathcal{U}$ for $i\in \{0, \ldots, T\}$ and some $T\in \mathbb{N}$.  
At times we will also consider trajectories in state space only, in which case a trajectory corresponds
to a sequence of states $\tau = (\bx_0,....,\bx_T)$ with $\bx_i\in\mathcal{X}$ for $i\in \{0, \ldots, T\}$.

@@

Dagger has been experimentally shown to perform  well in both simulation and real world experiments and furthermore
has guarantees guarantees the convergence to the supervisor's policy {\color{blue} [-- F: ONLY AN ERROR BOUND, RIGHT? --]}. 
However, policy rollouts can result in the robot entering dangerous states. For example, when DAgger was used on a
quad-copter the robot repeatedly crashed into trees before successfully learning to avoid them \cite{ross2013learning}.
We work under the assumption that a state can be risky for 2 reasons: 1) it lies in an area with a low density of
previously trained states, which potentially means our policy will mis-predict the supervisor and incur high surrogate
loss \cite{tokdar2010importance} 2) the surrogate loss of the current policy at the current state is high, so the state, although visited, does
not model the supervisor's control inputs correctly. Furthermore providing correct control inputs or ``labeling'' at each
iteration and each state in DAgger can be tedious or costly.

@@

\subsection{Estimation of Quantile Level Sets}\label{sec:level}
{\color{blue} [-- REVISED UP TO HERE --]}
Support estimation can be defined as minimizing a function that contains the volume of a given probability distribution. Formally we define it as follows let $\mathbf{x}_1,...,\mathbf{x}_n $ be i.i.d. random variables in a set $\mathcal{X}$ with distribution $P$. Let $\mathcal{G}$ be a class of measurable subsets of $\mathcal{X}$ and let $\lambda$ be a real function defined on $\mathcal{G}$. The quantile function with respect to $(P,\lambda,\mathcal{G})$ is

@@

\begin{figure}[t!]
\centering
\includegraphics[width=6cm, height=6cm]{figures/race_car.png}
\caption{ An example image from the racing simulator of a sampled polygonal track.. The car is meant to drive around an not fall off. It is placed back on the center if it does fall off. The supervisor drives around the track for 2 iterations and HOG features are extracted from the images to form a feature space of }

\vspace*{-10pt}
\label{fig:race_car}
\end{figure}

\begin{figure}[t!]
\centering
\includegraphics[width=8cm, height=6cm]{figures/dagger_shiv_one_class.eps}
\caption{We compare performance in terms of minimization of the underlying cost function $c(\bx,\bu)$, which is the  number of times the car left the track versus the number of queries made to the supervisor. In Fig. \ref{fig:car_cost}, we plot the performance of DAgger, SHIV and SHIV using only the traditional One-Class SVM without the modification.  Initial results, which are run for 6 iterations each and are averaged over 40 levels, shown in Fig. \ref{fig:car_cost} suggest an $70\%$ reduction in the number of queries needed for SHIV compared to DAgger,}
\vspace*{-10pt}
\label{fig:car_cost}
\end{figure}

@@

\begin{figure}[t!]
\centering
\includegraphics[width=6cm, height=6cm]{figures/box2D_arm.png}
\caption{  The arm, modeled with Box2D, consists of 3 degrees of freedom and has a parallel jaw gripper. For input the human demonstrator provides controls through an XBox controller. The left and right joystick determined the x and y velocity of the center of the end effector. The left bumper of the controller provided a binary input for opening and closing the gripper. Our state $\bx$ consisted of the 3 dimensional pose of the six objects on the table (translation and rotation), the 3 joint angles of the arm and a scalar value in the range $[0,1]$ that measured the position of the gripper, 1 being opened and zero being closed. }

\vspace*{-10pt}
\label{fig:box2D_arm.png}
\end{figure}

\subsection{Surigcal Experiment}

\section{Discussion and Future Work}

>>>>

\begin{abstract}
The SHIV algorithm proposes an alternative to previous query selection methods for stream-based active learning to the non-stationary distributions encountered in robot control.  To reduce operator burden, the Svm-based reduction in Human InterVention (SHIV) algorithm only requests supervision for sufficiently risky or novel states.

In traditional imitation learning, a human expert provides a few demonstrations to the robot, which then \emph{offline} learns a policy to extrapolate these examples to new situations. One challenge is that during execution of the learned policy, a small but inevitable error will amplify, taking the robot further and further away from the regions of state space for which it has seen examples.

\emph{Online} imitation learning algorithms such as DAgger address this challenge by iteratively executing a trained policy, and asking the expert for additional examples along the states visited by that policy. Such algorithms perform well in new situations, but rely much more heavily on extensive human supervision.

Our goal is to reduce the amount of human supervision required in imitation learning, without sacrificing performance. \emph{Our key insight is that the robot should ask for supervision only when its current trained policy has high risk.}

We introduce a method for online imitation learning that evaluates risk based on novelty of a state, as well as prior difficulty in labeling a nearby states. Our algorithm, SHIV (SVM-Based Reduction in Human Intervention), follows the trained policy when evaluated risk is low, and asks for human supervision otherwise. Our experiments with a toy car domain, as well as with non-prehensile grasping in clutter, suggest that our method can better utilize a human supervision budget, achieving higher success rates for the same number of queries to the expert.

@@

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/teaser.pdf}
\caption{ 
Given a set of states corresponding to images of a race-track simulation (visualized in $\mathbb{R}^2$) with expert
control inputs, we consider a quantile super levelset estimate (shaded) as an approximation of the riskiness of
executing the trained policy at a given state. By varying the quantile threshold, we can increase or decrease the size
of the quantile super-levelset area. In our approach, points outside the super-levelset are considered risky and, during
training, roll-outs of the policy that leave the super-levelset require control demonstrations from the supervisor,
while roll-outs within the super-levelset can be performed autonomously. 
}
\vspace*{-10pt}
\label{fig:dis_traveled}
\end{figure}

@@

\section{Problem Statement}
\subsection{Notation and Background}
We denote by $\mathcal{X}$ the set consisting of \emph{observable states} for a robot task, consisting, for example of
high-dimensional vectors corresponding to images from a camera, or joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of \emph{allowable control inputs} for the robot, which can be discrete or continuous in
nature. The environment is
modeled to have dynamics that are stationary and Markovian, such that the the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in \mathcal{U}$,
so that $p(\bx_{N+1}|\bu_{N},\bx_{N}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{N+1}|\bu_{N}, \bx_N)$.  
When no confusion arises, we denote the probability density over the initial state also by $p:\mathcal{X}\to
\mathbb{R}$. By a trajectory, we mean a finite series of pairs of $T$ states visited and corresponding
control inputs at these states $\tau = (\mathbf{x}_0,\mathbf{u}_1, ...., \mathbf{x}_T,\mathbf{u}_T)$, where $\bx_i\in \mathcal{X}$
and $\bu_i\in \mathcal{U}$ for $i\in \{0, \ldots, T\}$ and some $T\in \mathbb{N}$.  
At times we will also consider trajectories in state space only, in which case a trajectory corresponds
to a sequence of states $\tau = (\bx_0,....,\bx_T)$ with $\bx_i\in\mathcal{X}$ for $i\in \{0, \ldots, T\}$.

@@

Dagger has been experimentally shown to perform  well in both simulation and real world experiments and furthermore under the assumption of the policy being no regret has theoretical properties of being linearly in time horizon T to the supervisor's policy. 
However, policy rollouts can result in the robot entering dangerous states. For example, when DAgger was used on a
quad-copter the robot repeatedly crashed into trees before successfully learning to avoid them \cite{ross2013learning}.
We work under the assumption that a state can be risky for 2 reasons: 1) it lies in an area with a low density of
previously trained states, which potentially means our policy will mis-predict the supervisor and incur high surrogate
loss \cite{tokdar2010importance} 2) the surrogate loss of the current policy at the current state is high, so the state, although visited, does
not model the supervisor's control inputs correctly. Furthermore providing correct control inputs or ``labeling'' at each
iteration and each state in DAgger can be tedious or costly.

@@

\subsection{Estimation of Quantile Level Sets}\label{sec:level}
Support estimation can be defined as minimizing a function that contains the volume of a given probability distribution. Formally we define it as follows let $\mathbf{x}_1,...,\mathbf{x}_n $ be i.i.d. random variables in a set $\mathcal{X}$ with distribution $P$. Let $\mathcal{G}$ be a class of measurable subsets of $\mathcal{X}$ and let $\lambda$ be a real function defined on $\mathcal{G}$. The quantile function with respect to $(P,\lambda,\mathcal{G})$ is

@@

\begin{figure}[t!]
\centering
\includegraphics[width=8cm, height=6cm]{figures/dagger_shiv_one_class.eps}
\caption{We compare performance in terms of minimization of the underlying cost function $c(\bx,\bu)$, which is the  number of times the car left the track versus the number of queries made to the supervisor. In Fig. \ref{fig:car_cost}, we plot the performance of DAgger and SHIV.  Initial results, which are run for 6 iterations each and are averaged over 40 levels, shown in Fig. \ref{fig:car_cost} suggest an $70\%$ reduction in the number of queries needed for SHIV compared to DAgger,}
\vspace*{-10pt}
\label{fig:car_cost}
\end{figure}

@@

\subsection{Surgical Experiment}
\todo{Need help discussing this better}
In this experiment, we look at the task of inserting a needle into tissue in preparation for suturing. Needle insertion can be a challenging task for surgeons and learning a controller to do so could alleviate some of the burden placed on surgeons. We collected data from a surgeon, Dr. Douglas Boyd, on the Da Vince Research Kit.

Dr. Boyd demonstrated a series of trajectories that started at an initial attempted needle insertion pose and applied the neccesary corrections to achieve a correct pose. We used three of these demonstrations as our initial dataset $\mathcal{D}_0$. Dr. Boyd couldn't be around for the interactive labeling of both SHIV and DAgger, so we had to hardcore and expert controller that computed the transformation between the desired insertion pose and the current pose. We did this by calculating the inverse of a transformation matrix.

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth, height=6cm]{figures/needle_insertion_results.eps}
\caption{We compare performance in terms of minimization of the underlying cost function $c(\bx,\bu)$, which is euclidean distance between translation in centimeters. In Fig. \ref{fig:car_cost}, we plot the performance of DAgger, SHIV and SHIV using only the traditional One-Class SVM without the modification.  Initial results, which are run for 20 iterations each and are averaged over 40 different initial starting positions, shown in Fig. \ref{fig:car_cost} suggest an $67\%$ reduction in the number of queries needed for SHIV compared to DAgger,}
\vspace*{-10pt}
\label{fig:car_cost}
\end{figure}

\section{Discussions and Future Work}
SHIV currently provides a way to apply active learning to online imitation learning, however its use of level set estimation of the underlying distribution introduces several hyper parameters. As Fig. shown these can significantly effect performance if not tuned correctly. Future work will look at using cross validation on a hold out set of data to more effectively tune parameters.

Another potential issue is that each iteration of SHIV requires solving a quadratic program with the Gram matrix. As we scale to harder tasks that require significantly more data, this can become problematic. Future work will look at using techniques like PCA and random feature projections to scale efficiently \todo{cite something}.

Lastly, our approach is promising in simulation. However we need to evaluate how our active learning method will work on a real robot with noisy sensor observations. Future work will be doing the grasping in clutter experiments on an actual robot.

\section{Acknowledgments} 
This work is supported in part by the U.S. National Science Foundation under Award IIS-1227536 and NSF-Graduate Research Fellowship.. 
We thank UC Berkeley and our colleagues who gave feedback and suggestions, in particular Sanjay Krishnan, Siddarth Sen, Steve McKinley and Sergey Levine.
